---
title: "3/30/20 notes edld 610"
author: "JP"
date: "3/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Notes

## Introduce Yourself
1. JP
2. I wanted to take it last year but I wanted to finish my required course material.
3. Slowly losing it, but started baking bread, making pizza, and cooking
4. furry friend is lucy

## Notes

## Basic Data Types

### Vectors

- whole number is an integer, while doubles are numbers that can be between integers (floating, 6.47)

### function to create vector
- `c` is the function to create vectors

### list not a vector
- list is a vector

### difference between matrix and data frame
- matrix is a vector with dimension attribute, everything has to be of the same type, everything is double
- data frame, every column could be of a different type (integer, double, character, etc)

### coercion mean
- implicit and explicit coercion
- you have a vector, with characters and integers
- if this way, it will be coerced into the simplier type, a single type

- changing `is.numeric` may not be best to `is.integer` or `is.double`
- most times it will be okay with just `is.numeric`

--- 
## Coercion
- character in a vector, everything will be implicitly coerced into characters

### Explicit coercion
- can make whatever vector you want

---
## checking types
`typeof` function to see what type of vector each are

---
## piping
```{r}
library(magrittr)

typeof(as.integer(c(7L, 3.25)))

c(7L, 3.25) %>% 
  as.integer() %>% 
  typeof()
```

