---
title: "4_13_20 notes"
author: "JP"
date: "4/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 4/13/20 notes for EDLD 610 - Func Prog

```{r}
install.packages('patchwork')
```

patchwork is a package that can be used to organize plots on a slide/page.

Tidyeval is created to try and help with writing functions. However, it is still more difficult than base r.

```{r}
library(tidyverse)
```

# Purr

purr will return a list using `map` functional.

map is the equivalent of using lapply

using `~` is usful for shortcuts in purrr, it removes the `function(x)` in code

rather than `x`, you'd use `.x`

# Purr Variants

`map_dbl`, `map_int`, `map_char`, `map_lgl`

these will return the atomic vectors described in each variant

map will always return a list though

## Activity

```{r}
econ <- economics %>% 
  mutate(year = lubridate::year(date))

econ

by_year <- split(econ, econ$year)

str(by_year)

names(economics)
```

```{r model building}
# pce --> unemployment percentage over time

# don't need the code below because you don't need to have this prior to looping since this is using map. This is for for functions.
# year_finding <- rep(NA, length(by_year))

year_per <- map(by_year, ~mutate(.x, percent = unemploy/pop))


# models for each year
model <- map(year_per, ~lm(percent ~ pce, data = .x))

str(model)


#getting coefficients for each year
coef_mod <- map(model, ~coef(.x))
# same thing
# coef_mod <- map(model, coef)

str(coef_mod)

coef_mod[c(1:2, length(coef_mod))]




# extracting slopes
pce_only <- map(coef_mod, 'pce')
# if you want a single vector, pce is all double
pce_only <- map_dbl(coef_mod, 'pce')

pce_only
```

**Remember to put in the `~` as this is the equivalent of using the function(x)**

piping can be used for purr functions




`map_lbl` can be used to examine columns that are only of a certain atomic vector, in this case, only looking at numeric columns. 
```{r}
map_lgl(iris, is.numeric)
```





