---
title: "Midterm"
author: "JP"
date: "5/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)

# install.packages('glue')
```

# EDLD 610 - Functional Programming Midterm

```{r}
download_file <- function(year) {
    link <- glue::glue("https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx")
    rio::import(link, setclass = "tibble", na = c("-", "--", "*"))
}

```

1. (10 points) Use the function above to download all the data from 2014-15 to each of the past 4 school years and bind it into a single data frame, using a single function (i.e., one line of code). Note, this may take a minute or two to run, depending on your Internet speed.

```{r}
library(tidyverse)

# need to think through how to do this in a programming way

data <- map_df(c(15, 16, 17, 18), 
               ~download_file(.x)) %>% 
  janitor::clean_names()

```

Conduct some basic data cleaning to make your data file look like the following.

* Filter for only student groups coded as "White" or "Hispanic/Latino".
* Select variables related to the number of students in each of the levels (1:4), and not percentages or collapsed levels.
* Remove any row that has missing data in any of the n variables


```{r}
names(data)

data$student_group <- str_to_lower(data$student_group)

data_sub <- data %>% 
  filter(student_group == 'white' |
           student_group == 'hispanic/latino') %>% 
  dplyr::select(academic_year,
                district,
                school,
                student_group,
                grade_level,
                number_level_4,
                number_level_3,
                number_level_2,
                number_level_1) 

data_long <- data_sub %>% 
  pivot_longer(number_level_4:number_level_1, 
               names_to = 'level',
               values_to = 'n') %>% 
  drop_na(n)

  
```

2. (10 points) Collapse (sum) the n across grades and school for each district within each academic year, student group, and level. At the same time, calulate the number of schools (unique schools listed) for each district. Also move your data to a wider format where the student groups become new columns, filled by the district n. Remove districts that did not report numbers for one or both student groups (i.e., drop missing data from the new columns). Your data frame should look like the below

```{r}
# n_variables <- data_long %>%
#   group_by(district, academic_year, student_group, level) %>%
#   nest() %>%
#   mutate(n_students = map_dbl(data, ~sum(.x$n))) %>%
#   unnest() %>%
#   select(academic_year, district, student_group, n_students, school, level) %>%
#   group_by(district) %>%
#   mutate(n_schools = n_distinct(school)) %>% 
#   ungroup()

n_students <- data_long %>%
  group_by(district, academic_year, student_group, level) %>%
  mutate(students = sum(n)) %>%
  ungroup() %>% 
  select(academic_year, district, student_group, students, level) %>%
  distinct()

n_schools <- data_long %>%
  dplyr::select(district, school) %>%
  distinct() %>%
   group_by(district) %>%
   summarize(schools = n()) %>% 
  ungroup()

joined <- left_join(n_students, n_schools)

wider <- joined %>%
  pivot_wider(names_from = student_group,
              values_from = students) %>% 
  drop_na() %>% 
  arrange(district)

```

```{r}
library(gapr)
wider[1:4, ] %>% # data source - just first four rows for first district
  estimate_v("white", "hispanic/latino")
```

1. (10 points) Estimate an achievement gap effect size for every school district in the state that reported data on both student groups (i.e., using the data we created above), for each academic year. Note, on my computer this took about 38 seconds. It will likely take a bit longer for you, but shouldn’t take too long. You may get a warning. Just go ahead and ignore that for now.

```{r}
eth_diff <- wider %>%
  group_by(academic_year, district) %>% 
	nest() %>%
	mutate(gap_es = map(data, ~estimate_v(.x, "white", "hispanic/latino")))

v <- eth_diff %>% 
  dplyr::select(-data) %>% 
  unnest()

```


2. (10 points) The plot below shows the achievement gap estimate for one school district across years. Produce a similar plot to the below for each of the first 100 school districts. Make sure you don’t miss the caption noting the number of schools in the district.


```{r}
dists <- unique(v$district)[1:100]

plots <- v %>% 
  as_tibble() %>% 
  filter(district %in% dists)
```

```{r}
bandon_plot <- plots %>% 
  filter(district == 'Bandon SD 54') %>% 
  ggplot(aes(as.factor(academic_year), v)) +
  geom_line(color = 'gray70') +
  geom_errorbar(aes(ymin = v - v_se, 
                    ymax = v + v_se)) +
  geom_hline(yintercept = 0,
             color = "red") +
  geom_point(color = 'dodgerblue') +
  theme_minimal() +
  labs(x = 'School Year',
       y = 'Average Achievement Difference (SD units)',
       title = 'Achievement Gap Estimates: Bandon School District',
       subtitle = 'Negative values indicate Hispanic/Latino students have higher average achievement\nPositive values indicate White students have higher average achievement',
       caption = 'Note: District included 3 schools') 

bandon_plot
```

```{r}

all_plots <- left_join(plots, n_schools) %>% 
  mutate(label_gap_est = glue::glue('Achievement Gap Estimates: {district}'),
         label_note = glue::glue('Note: District included {schools} schools.')) %>% 
           group_by(district, label_gap_est, label_note) %>% 
           nest() %>% 
  mutate(plots = pmap(list(district, label_gap_est, label_note, data),
                      ~ggplot(..4, aes(as.factor(academic_year), v)) + geom_line(color = 'gray70') +
  geom_errorbar(aes(ymin = v - v_se, 
                    ymax = v + v_se)) +
  geom_hline(yintercept = 0,
             color = "red") +
  geom_point(color = 'dodgerblue') +
  theme_minimal() +
  labs(x = 'School Year',
       y = 'Average Achievement Difference (SD units)',
       title = ..2,
       subtitle = 'Negative values indicate Hispanic/Latino students have higher average achievement\nPositive values indicate White students have higher average achievement',
       caption = ..3)))
                        
all_plots$plots[[1]]
all_plots$plots[[40]]
```

3. (10 points) Save the plots into a “plots” directory. Make sure the file names are meaningful.

```{r}
dir.create(here::here('plots', 'districts'))

files <- str_replace_all(tolower(all_plots$district), " ", "-")

paths <- here::here('plots', 'districts', glue::glue("{files}.jpeg"))

paths

walk2(paths, all_plots$plots, ggsave,
      width = 9.5, 
      height = 6.5,
      dpi = 500)
```