---
title: "Lab 4"
author: "JP"
date: "5/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Writing a summary function
Write a function that takes a given data frame as its first argument, a column to summarize as its second argument, and any set of other columns to summary operations by as the third argument.

Function behavior
By default, the function should report:

Total number of observations
Number of non-missing observations
Number of missing observations
Mean
Standard deviation
Minimum
Maximum
However, the function should be flexible so it can take any function(s) as its input to summarize by, provided the function results in a single value (e.g., anything you could feed to dplyr::summarize). You’ll need an additional optional argument (should be the last argument in your function) that lists the function(s) to use in the summary. Hint: You’ll want to have this argument pre-populated with default functions.

Errors
Embed an error if

Anything other than a data frame is supplied as the first argument
The provided column to summarize is not numeric.
Warnings
Embed a warning if

Any of the grouping variable(s) are numeric (generally you would expect the grouping variable to be a categorical variable)
Some guidance
If you want to use {dplyr} you’ll need to use non-standard evaluation (NSE), so either something like !!enquo(group_var), where group_var is the formal argument of your function, or {{group_var}}. Also, dplyr::pull (together with NSE) will likely be helpful for checking types for warnings/errors.

Alternatively, you might consider spliting the dataset by the grouping variable, then looping over each split to apply the function(s) supplied.

```{r}
library(tidyverse)

grouped_summarized <- function(df, 
                               group_var, 
                               summarized_var, 
                               functions = list(n = ~length(.),
                                            valid_cases = ~sum(!is.na(.)),
                                            missing_cases = ~sum(is.na(.)),
                                            mean_var = ~mean(.),
                                            sd_var = ~sd(.),
                                            min_var = ~min(.),
                                            max_var = ~max(.))
                               ){
  if(!is.data.frame(df)) {
        stop("Data must be in the form of a data frame.")
    }
    if(!is.numeric(pull(df, {{summarized_var}}))) {
        stop("Summarized column needs to be numeric.")
    }
    if(is.numeric(pull(df, {{group_var}}))) {
        warning("The grouping variable is numeric. Change the numeric value to be a factor.")
    }
  
  df %>% 
    group_by({{group_var}}) %>% 
    summarize_all(functions)
}

iris %>% 
  grouped_summarized(Species, Sepal.Width)

airquality %>%
    mutate(Month = as.factor(Month)) %>%
    grouped_summarized(Month, Ozone)

```




