---
title: "lab 1"
author: "JP"
date: "4/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 1

```{r}
library(tidyverse)

set.seed(123) # Set random number generation

m <- lm(mpg ~ hp + cyl + drat, mtcars)

l <- list(a = list(m = matrix(1:12, ncol = 3),
                   v = 1:7),
          b = data.frame(student = 1:15,
                         score = rnorm(15, 100, 10)))
```

Run the above code

1. From the model results:

- Extract the qr tolerance (tol).
- Extract the term labels from the model.
* hint: Use str to inspect the list. This one is tricky and involves pulling an attribute from an attribute.

```{r}
attributes(m)

str(m)

m$qr$tol

attributes(m$model)

terms <- attr(m$model, 'terms')

attr(terms, 'term.labels')
```


2. From the list

- Use at least two different methods to extract m.
- Extract the third column of m. Maintain the dimensions (matrix structure).
- Extract the score for student 7.

```{r}
l$a$m
l$a['m']
l[['a']]['m']


l[['a']][['m']][, 3]

m <- matrix(1:12, ncol = 3)
m[, 3]


l[['b']][7, 2]

l$b[7, 2]

```


## Loops

3. Use the “airquality” dataset, which is part of base R. Split it by Month
```{r}
airquality <- airquality %>% 
  janitor::clean_names()

by_month <- split(airquality, airquality$month)

by_month

typeof(by_month)
```



4. Use a for loop to calculate the mean Ozone by month (remember to remove missing data from the calculation)
```{r}

result <- rep(NA, length(by_month))

for(i in seq_along(by_month)) {
  result[i] <- mean(by_month[[i]]$ozone, na.rm = TRUE)
}

result

```


5. Replicate this calculation with lapply, sapply, and vapply
```{r}
# use anonymous function

l_mean <- lapply(by_month, function(x) mean(x$ozone))
l_mean
```



6. Produce separate plots showing the relation between Ozone and Solar.R for each month (with a single loop).

7. Use a for loop to save the plots to a folder on your computer